### 2021-05-16 R/ARR.Rmd

---
title: "New Effect Sizes: Acclimation Response Ratio (ARR)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ARR (Acclimation Response Ratio)

### Point estimate and definition

$$
\text{ARR} = \frac{ \text{CT}_{\text{max}[T_2]} - \text{CT}_{\text{max}[T_1]}}{T_2 - T_1},
$$ where *T* stands for temperature in Celsius and $T_2 > T_1$.

### Sampling variance for AAR

#### Independent case

$$
s^2(\text{AAR}) = \left( \frac{1} {T_2 - T_1} \right)^2 
\left( \frac{sd^2_{[T_1]}} {n_{[T_1]}} + \frac{sd^2_{[T_2]}} {n_{[T_2]}} \right).
$$

#### Dependent case

$$
s^2(\text{AAR}) = \left( \frac{1} {T_2 - T_1} \right)^2 
\left( \frac{sd^2_{[T_1]} + sd^2_{[T_2]} - r_{[T_1T_2]}sd_{[T_1]}sd_{[T_2]}} {n} 
\right).
$$

### Basics about variance

See <https://en.wikipedia.org/wiki/Variance>

Adding a constant ($a$) does not change the variance of a random variable $X$. $$
\operatorname{Var}(X+a)=\operatorname{Var}(X)
$$

Multiplying a constant ($a$) does increase the variance of $X$ by $a^2$.

$$
\operatorname{Var}(aX)=a^2\operatorname{Var}(X)
$$

When we consider two random variables $X$ and $Y$, we need to know their covariance.

$$
\operatorname{Var}(aX\pm bY)=a^2\operatorname{Var}(X)+b^2\operatorname{Var}(Y) \pm 2ab\operatorname{Cov}(X,Y)
$$