### 2021-06-03 MS/supp.Rmd
Updated supp to include details on searches

---
title: "Supplemental Materials for: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Potter, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: yes
    toc_depth: 6
    toc_float: yes
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
editor_options: 
  chunk_output_type: console
---

```{r loadpackages, echo = FALSE, results = 'hide'}
pacman::p_load(png, jpeg, cowplot, flextable, tidyverse)
```

# Supplementary Materials for Survey Study Selection

Here we provide greater detail on the search strategy, keyword selection and the numbers of final studies included at each step pf the screening process for our survey. Figure \@ref(fig:figS1) outlines our PRSIMA flowchart detailing the number of papers hit, their sources and how many records were excluded and included at each screening phase. The full search strings used to refine our search query can be found in Table \@ref(tab:tabS1). Note that results from our final searches (25 & 28) are the ones we used. Our query was restricted to a pre-specified set of comparative physiology journals within the last 6 years of publication. The journals used and their ISSN numbers are provided in Table \@ref(tab:tabS2). In total, out of the 79 papers screened at the full-text stage we ended up including 62 full texts. We excluded N = 17 meta-/comparative analyses because they: 1) did not include physiological traits (N = 4); 2) were empirical studies that collected their own data and did not collate data from the literature (N = 10); 3) Were not actual quantitative syntheses, but rather systematic reviews (N = 1) and 4) were studies with a focus on biomechanics (N = 2).

```{r figS1, fig.align='center', echo = FALSE, fig.cap="PRISMA flowchart for study selection and inclusion"}

img1 <- readPNG("../figs/PRISMA_diagram_v0.png")

p1 <- cowplot::ggdraw()+
        cowplot::draw_image(img1)
p1
```

```{r tabS1,  echo = FALSE, tab.cap="Log of search string development"}

table <- read.csv("../tables/Table_S1.csv")

table_s1 <- flextable(table)                                                                            %>% 
            flextable::compose(j = 1, part = "header", value = as_paragraph(as_b("Date")))              %>% 
            flextable::compose(j = 2, part = "header", value = as_paragraph(as_b("Search Number")))     %>% 
            flextable::compose(j = 3, part = "header", value = as_paragraph(as_b("Search String")))     %>% 
            flextable::compose(j = 4, part = "header", value = as_paragraph(as_b("Number of Records"))) %>% 
            flextable::compose(j = 5, part = "header", value = as_paragraph(as_b("Comments")))          %>% 
            flextable::compose(j = 6, part = "header", value = as_paragraph(as_b("Notes"))) 
 table_s1
```

```{r tabS2, echo = FALSE, tab.cap="Log of search string development"}

table <- read.csv("../tables/Table_S2.csv")

table_s2 <- flextable(table)                                                                            %>% 
            flextable::compose(j = 1, part = "header", value = as_paragraph(as_b("Journal Name")))      %>% 
            flextable::compose(j = 2, part = "header", value = as_paragraph(as_b("ISSN Number")))       %>% 
            flextable::width(j = 1, 5)
table_s2
```