### 2021-07-22 R/survey_data_results.Rmd
tidying formatting of survey results

---
title: "Survey Results from: Meta-analytic approaches and new effect sizes to account for heterogeneous effects in comparative physiology"
author: Daniel W.A. Noble, Patrice Pottier, Sammy Burke, Szymek Drobniak, Malgorzata Lagisz, Rose O’Dea, Shinichi Nakagawa
date: "`r Sys.Date()`"
output:
  html_document:
    theme: paper  
    highlight: tango  
    toc: true
    toc_depth: 5
    number_sections: false  
    toc_float:
      collapsed: false
      smooth_scroll: false
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

pacman::p_load(dplyr,ggplot2,here, pander, janitor, magrittr,ggalluvial,ggpubr)

# loading data
raw <- read.csv(here("data","MA_CompPysio_Survey.csv"))
metadata <- read.csv(here("data","MA_CompPysio_Survey_metadata.csv"))

# tidying data
dat <- raw
names(dat) <- metadata$Procesed.Column.Name[match(names(dat), metadata$Raw.Column.Name)]

# tidying categorical options
dat$meta_analysis_self_described <- factor(gsub(" \\(.*", "", dat$meta_analysis_self_described), levels = c("yes","no","unclear/other"))
dat$comparative_physiology_main_focus <- factor(gsub(" \\(.*", "", dat$comparative_physiology_main_focus), levels = c("yes","no","unclear/other"))

# fixing format of journals

#dat$DOI[dat$journal=="2017"] # Typo
dat$journal[dat$journal=="2017"] <- "Biology Letters"

#dat$DOI[grep("Proceedings", dat$journal)] # all the same journal
dat$journal[grep("Proceedings", dat$journal)] <- "Proceedings of the Royal Society B: Biological Sciences"

# standardising DOI format
dat$DOI <- gsub(".*org/", "", dat$DOI)
dat$DOI <-gsub(".* ", "", dat$DOI)

# ggplot2 theme for plots
background.colour <- "white"
text.colour <- "black"
# Font Choice
#loadfonts()
text.font <- "Helvetica"
line.width <- 0.6

tm <- theme(panel.background = element_rect(fill = background.colour),
            plot.background = element_rect(fill = background.colour),
            panel.grid = element_blank(),
            text =  element_text(family = text.font, colour = text.colour),
            axis.text = element_text(colour = text.colour),
            axis.title.x = element_text(margin = margin(10,0,0,0)),
            axis.title.y = element_text(margin = margin(0,10,0,0)),
            axis.line = element_line(colour = text.colour, size = line.width),,
            axis.ticks = element_line(colour = text.colour, size = line.width),
            plot.title = element_text(hjust = line.width),
            legend.background = element_rect(fill = background.colour),
            legend.key = element_rect(fill = background.colour),
            plot.margin = margin(6,6,6,6))

tm2 <- theme(panel.background = element_rect(fill = background.colour),
            plot.background = element_rect(fill = background.colour),
            panel.grid = element_blank(),
            text =  element_text(family = text.font, colour = text.colour, size = 15),
            axis.text = element_text(colour = text.colour),
            axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5),
            axis.title.x = element_text(margin = margin(10,0,0,0)),
            axis.title.y = element_text(margin = margin(0,10,0,0)),
            axis.line = element_line(colour = text.colour, size = line.width),,
            axis.ticks = element_line(colour = text.colour, size = line.width),
            plot.title = element_text(hjust = line.width),
            legend.background = element_rect(fill = background.colour),
            legend.key = element_rect(fill = background.colour),
            plot.margin = margin(15,15,15,15))
```

## Sample of `r nrow(dat)` papers {.tabset}

### Publication years {.tabset .tabset-fade .tabset-pills}

```{r}
dat %>% dplyr::group_by(pub_year) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = pub_year, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,13,2)) +
  scale_x_continuous(breaks = min(dat$pub_year):max(dat$pub_year)) +
  labs(x = "publication year",
       y = "number of papers")
```

### Meta-analysis {.tabset .tabset-fade .tabset-pills}

Answer to the question "Is paper self-describing as containing a meta-analysis (or meta-regression) or actually using meta-analysis/regression approach?".

```{r}
dat %>% dplyr::group_by(meta_analysis_self_described) %>% dplyr::tally() %>% 
  ggplot() + tm +
  geom_col(aes(x = meta_analysis_self_described, y = n),
           width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  labs(x = "meta-analysis or meta-regression",
       y = "number of papers") 
```

### Comparative physiology {.tabset .tabset-fade .tabset-pills}

Answer to the question "Is the study's main focus on comparative physiology?"

- yes = focused only on physiological traits very clearly)  
- no =  does have physiological traits, but is interested in broader questions involving many different traits

```{r}
dat %>% 
  ggplot() + tm +
  geom_bar(aes(x = comparative_physiology_main_focus,
               fill = meta_analysis_self_described,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_viridis_d() +
  labs(x = "main focus on comparative physiology",
       y = "number of papers",
       fill = "meta-analysis") 
```

## Meta-analytic models {.tabset}

Note that single papers can report more than one approach (e.g., use different effect sizes and different models).

### Effect sizes used {.tabset .tabset-fade .tabset-pills}

```{r}
dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type) %>%
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Means = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(prop) %>% mutate(effect=factor(effect,levels=unique(effect))) %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=10)) +
  geom_col(aes(x = effect, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,20,2)) +
  labs(x = "Effect Size",
       y = "Number of Papers") +
  coord_flip()
```

### Statistical models used {.tabset .tabset-fade .tabset-pills}

```{r}
summary <-  dat %>% dplyr::group_by(stat_model_type) %>% janitor::tabyl(stat_model_type) %>%
  dplyr::summarise(`Multilevel Meta-analysis Model` = dplyr::if_else(grepl("multilevel meta-analysis", stat_model_type), n, 0),
                  `Multilevel Metaregression Model` = dplyr::if_else(grepl("meta-regression models", stat_model_type), n, 0),
                    `Linear Regression` = dplyr::if_else(grepl("linear regression", stat_model_type), n, 0),
                 Other = dplyr::if_else(grepl("unclear/other", stat_model_type), n, 0),
                 `Vote Counting` = dplyr::if_else(grepl("vote counting", stat_model_type), n, 0),
                 `Weighted Regression` = dplyr::if_else(grepl("weighted regression including mixed effects models", stat_model_type), n, 0),
                 `Random Effects Model` = dplyr::if_else(grepl("random-effects meta-analysis", stat_model_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = (n / sum(n))) %>% arrange(prop) %>% mutate(effect=factor(effect,levels=unique(effect)))
rownames(summary) <- NULL

summary %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=10)) +
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),limits = c(0,0.4)) +
  labs(title = "Analytical Approach",
       y = "Frequency") +
  coord_flip()
```

### Weighted models? {.tabset .tabset-fade .tabset-pills}

Were effect sizes used in the analysis weighted by sampling variance?

```{r}
dat %>%  janitor::tabyl(meta_analysis_weighted) %>% 
  ggplot() + tm +  theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=10)) +
  geom_col(aes(x = meta_analysis_weighted,  y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(y = "Frequency") +
  coord_flip()
```

How do the two categorisations (meta-analysis and weighting) combine?

```{r}
dat %>% 
  ggplot() + tm +
  geom_bar(aes(x = meta_analysis_self_described,
               fill = meta_analysis_weighted,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_fill_viridis_d() +
  labs(x = "Meta-analysis?",
       y = "Proportion",
       fill = "Weighted?") 
```

## Heterogeneity {.tabset}

### Effect size impacted by 'nuisance heterogeneity'? {.tabset .tabset-fade .tabset-pills}

Could effect size be impacted by an additional continuous predictor (e.g., temperature in thermal limit studies, dosage in hormone manipulation studies)?

If "yes", does the study investigate any continuous environmental or methodological factors as moderators?

```{r}
dat %>% mutate(cont_moderator_variable_modelled = if_else(cont_moderator_variable_modelled == "", "N/A", cont_moderator_variable_modelled)) %>% 
  ggplot() + tm +
  geom_bar(aes(x = factor(cont_moderator_variable,levels=c("yes","no","unclear")),
               fill = cont_moderator_variable_modelled,
               y = ..count../sum(..count..)), width = 0.5,
           colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_fill_viridis_d() +
  labs(x = "Nuisance Heterogenity?",
       y = "Proportion",
       fill = "Modelled Nuisance\n Heterogenity?") 
```

### Reported Heterogeneity? {.tabset .tabset-fade .tabset-pills}

Of the studies that are weighted how many actually reported measures of heterogeneity?

```{r}
dat %>%  filter(!heterogeneity_reported=="N/A (e.g. they conducted unweighted meta-analysis)") %>% mutate(heterogeneity_reported = if_else(heterogeneity_reported == "unclear/other (add comment)", "other", heterogeneity_reported)) %>% janitor::tabyl(heterogeneity_reported) %>% dplyr::arrange(-n) %>% 
  ggplot() + tm +
  geom_col(aes(x = factor(heterogeneity_reported,levels=heterogeneity_reported), y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  labs(x = "Heterogeneity reported?",
       y = "Frequency") 
```

## Non-independence and publication bias {.tabset}

### Handling sources of non-independence {.tabset .tabset-fade .tabset-pills}

Did the analysis account for - or mention - additional non-independence arising from, e.g., phylogenetic relatedness or shared control treatments? The actual way of dealing with non-independence (n-i) will vary from study to study.

```{r}
dat %>%  janitor::tabyl(nonindependence_mentioned) %>% dplyr::arrange(n) %>% 
  ggplot() + tm + theme(axis.title.y=element_blank(),
                        axis.text.y = element_text(size=10)) +
  geom_col(aes(x = factor(nonindependence_mentioned,levels=nonindependence_mentioned), y = percent), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,1)) +
  scale_x_discrete(labels = c('n-i dealt with w/o modelling', 'n-i modelled', 'unclear/other/not mentioned')) +
  labs(title = "Non-independence handling",
       y = "Frequency") +
  coord_flip()
```

### Publication bias {.tabset .tabset-fade .tabset-pills}

What type of publication bias analysis was used in the analysis? Multiple metrics can be used per paper, hence frequencies won't sum to 100%.

```{r}
dat %>% dplyr::group_by(pub_bias_type) %>% janitor::tabyl(pub_bias_type) %>%
  dplyr::summarise(GraphicalTest = dplyr::if_else(grepl("graphical tests", pub_bias_type), n, 0),
                  RegressionTest = dplyr::if_else(grepl("regression test", pub_bias_type), n, 0),
                    TrimAndFill = dplyr::if_else(grepl("trim-and-fill", pub_bias_type), n, 0),
                 TimeLag = dplyr::if_else(grepl("time-lag", pub_bias_type), n, 0),
                 OtherUnclear = dplyr::if_else(grepl("unclear", pub_bias_type), n, 0),
                 None = dplyr::if_else(grepl("no", pub_bias_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = (n / sum(n)*100)) %>% arrange(n) %>% mutate(effect = factor(effect,levels=unique(effect))) %>% 
  ggplot() + tm + theme(axis.title.y = element_blank(),
                        axis.text.y = element_text(size=10)) +
  geom_col(aes(x = effect, y = n), width = 0.5, fill = background.colour, colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0),breaks=seq(1,30,2)) +
  labs(x = "Effect Size",
       y = "Number of Papers") +
  coord_flip()
```

## Other plots {.tabset}

```{r}
plot1 <- dat %>% 
  ggplot() + tm2 + theme(legend.position = "top") +
  geom_bar(aes(x = comparative_physiology_main_focus,
               fill = cont_moderator_variable,
               y = ..count../sum(..count..)), width = 0.5,
           colour = NA, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Main focus on comparative physiology",
       y = "Number of papers",
       fill = "Continuous predictor present")
plot1

plot2 <- dat %>% 
  ggplot() + tm2 + theme(legend.position = "top") +
  geom_bar(aes(x = pub_year,
               fill = meta_analysis_self_described,
               y = ..count../sum(..count..)), width = 0.5,
           colour = NA, size = 0.5) +
  scale_y_continuous(expand=c(0,0)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = "Year of publication",
       y = "Number of papers",
       fill = "Meta-analysis")
plot2

plot3 <- dat %>%
  ggplot(mapping = aes(x = comparative_physiology_main_focus,
               y = meta_analysis_self_described)) + tm2 +
  theme(panel.border = element_rect(size = 0.5, fill = NA),
        panel.grid.major = element_line(colour = "grey92"),
        axis.text.y = element_text(angle = 90)) +
  geom_count(mapping = aes(size = ..n.., colour = "A")) +
  scale_size_continuous(range = c(20,40), guide = F) +
  scale_colour_brewer(palette = "Set2", guide = F) +
  # scale_y_continuous(expand=c(0,0)) +
  # scale_fill_viridis_d() +
  labs(x = "Main focus on comparative physiology",
       y = "Number of papers")
plot3

plot4 <- dat %>%
  ggplot(aes(axis1 = meta_analysis_self_described,
             axis2 = comparative_physiology_main_focus,
             axis3 = cont_moderator_variable)) +
  ggalluvial::geom_alluvium(aes(fill = comparative_physiology_main_focus), width = 1/4, alpha = 0.8) +
  tm2 + theme(axis.line = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position = "none", panel.grid = element_blank(), axis.text.y = element_blank()) +
  scale_fill_brewer(palette = "Set2") +
  ggalluvial::geom_stratum(fill = 'white', colour = 'black', width = 1/4) +
  geom_text(stat = 'stratum',
            aes(label = after_stat(stratum)),
            # label = c('weighted', 'unweighted', 'unclear'),
            colour = 'black') +
  scale_x_discrete(expand = c(.1, .1), limits = c("Meta-analysis", "Comparative physiol.", "Cont. moderator"))
plot4

panels <- ggarrange(ggarrange(plot1, plot2, nrow = 1),
                    ggarrange(plot3, plot4, nrow = 1, widths = c(1,2)),
                    nrow = 2, ncol = 1)
ggsave(panels, filename = "multipanel.pdf", device = "pdf")

```

## R Session Information   
```{r}
pander(sessionInfo(), locale = FALSE)
```