### 2021-07-22 MS/box.Rmd
Box updated with Losia's comments

---
bibliography: ../bib/refs.bib
csl: ../bib/the-journal-of-experimental-biology.csl
output: 
  bookdown::word_document2:
    toc: no
    toc_depth: 6
    number_sections: false
    reference_docx: template.docx
  bookdown::html_document2:
    code_folding: hide
    number_sections: no
    toc: no
    toc_depth: 6
    toc_float: yes
    bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## *BOX 1: Examples of how to derive sampling variances*

Here, we show how to obtain sampling variance for a slope (or a 'rate of change') when you have measurements of a physiological trait at two points along an environmental gradient (e.g., temperature, salinity, acidity). We then derive sampling variance for the difference between two slopes, and demonstrate how *SMD* can be corrected to account for differences in both units (e.g. cm or mg) and points on an environmental gradient (e.g. temperature of 24 and 30 \&degC). Finally, we introduce a useful approximation technique known as 'the delta method'. The sampling variances associated with these 'new' effect sizes below will allow comparative physiologists to take advantage of powerful meta-analytic models.

### Sampling variance for a slope between two points

Let's start with a real example that comparative physiologists can easily relate with, the slope of responses to temperature acclimation, or as we call it, the acclimation response ratio (*ARR*). *ARR* can be defined as a slope for acclimated physiological responses at two different temperature points, defined as:

$$
\text{ARR} = \frac{M_{1} - M_{2}}{T_2 - T_1},
$$

where *T* stands for temperature in degrees Celsius and $T_{2} > T_{1}$, and $M_{1}$ and $M_{2}$ are the average physiological responses (e.g., mean CTmax) at temperature points $T_{1}$ and $T_{2}$.

To obtain the sampling variance for this equation (slope), we first need to describe some basic properties of variance. Let's assume $M_{1}$ is a random variable, which means that this variable has a mean and standard deviation [note that this standard deviation is not the 'sample' but the 'sampling' standard deviation, which is often referred to as standard error - see Figure 1 in @Nakagawa2021b]. Multiplying a constant ($a$) will change the variance by the square of that constant ($a^2$) while adding or subtracting the constant ($b$) does not change the variance of $M_{1}$. This can be summarized as:

$$
\sigma^2(aM_1 \pm b)=a^2 * \sigma^2_{M_{1}}
$$

Also, when adding two random variables ($M_{1}$ and $M_{2}$), the combined variance is the sum of the variance of $X_{1}$ and the variance of $X_{2}$ plus 2 times the covariance between $M_{1}$ and $M_{2}$. This relationship can be written as:

$$
\sigma^2(M_1 \pm M_2) = \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cov}(M_1,M_2) 
= \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }
$$

where the co-variance $\text{Cov}(M_1,M_2)$ equals to the correlation multiplied by the square-root of the two variances $2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }$.

Importantly, when $M_{1}$ and $M_{2}$ are independent to each other their covariance is 0. In other words, if measurements are taken on two different groups of animals at two different temperatures ($T_1$ and $T_2$), then the covariances between these two sets of measurements are 0.

Therefore, when $M_{1}$ and $M_{2}$ are independent, we can obtain the sampling variance for *ARR* as:

$$
\sigma^2_{AAR} = \left( \frac{1} {T_2 - T_1} \right)^2 
\left( \frac{SD^2_{1}} {N_{1}} + \frac{SD^2_{2}} {N_{2}} \right)
$$

where $SD_{1}$ and $SD_{2}$ and $N_{1}$ and $N_{2}$ are standard deviations and sampling sizes at temperature $T_{1}$ and $T_{2}$. Readers may find it difficult to see how we obtained this equation. Let us explain further. The (sampling) standard error for $X_{1}$ is $SE_{1} = SD_{1} / \sqrt{N_{1}}$. Given this, the sampling variance for $X_{1}$ is $SD^2_{1} / N_{1}$, and the sampling variance for $X_{2}$ is $SD^2_{2} / N_{2}$. You can see those elements of the sampling variance of *ARR* in the formula above. The term $(1/(T_{2} - T_{1}))^2$ comes from recognizing $1/(T_{2} - T_{1})$ as a constant.

In ecological and evolutionary studies, it is not uncommon that we measure the same group of organisms at two temperatures. If we do, however, then we need to add the covariance between $M_{1}$ and $M_{2}$ ($\text{Cov}(M_{1},M_{2})$) in the equation. Note that, as above, the covariance equals the correlation multiplied by the square-root of two (sampling) variances. Therefore, the sampling variance of *ARR* can be now written as:

$$
\sigma^2_{AAR} = \left( \frac{1} {T_{2} - T_{1}} \right)^2 
\left( 
\frac{SD^2_{1}} {N_1} + \frac{SD^2_{2}} {N_2} - 2r \sqrt{\frac{SD^2_{1}} {N_1}  \frac{SD^2_{2}} {N_2}} 
\right)
$$

By assuming the number of organisms (*N*) are the same at the two temperatures, we can slightly simplify this formula:

$$ 
\sigma^2_{AAR} = \left( 
\frac{1} {T_2 - T_1} 
\right)^2
\left( 
\frac{SD^2_{1} + SD^2_{2}  - 2rSD_{1}SD_{2}} {N} 
\right)
$$

where *r* is the correlation between a set of measurements $M_1$ and $M_2$ from the same individuals at two points $T_1$ and $T_2$. As you may notice, we need the raw data to calculate *r*. Therefore, in reality, we often need to assume a certain value of *r*. When we do not have an estimate of *r* we can reasonably assume it to be 0.5 [@Noble2017].

### Sampling variance for the difference between two slopes

Now let us assume that we want to know the difference between two different *ARR* values: female *ARR*, $\text{ARR}_{f}$ and male *ARR*, $\text{ARR}_{m}$. Such a difference (*ARRD*) can be written as:

$$
\text{ARRD} = \text{ARR}_{f} - \text{ARR}_{m}.
$$

Using the properties of variance and the equations from above, the sampling variance for *ARRD* can be derived, when measurements at two temperature points are independent, as:

$$
\sigma^2_{ARRD} = \left( \frac{1} {T_2 - T_1} \right)^2 
\left( \frac{SD^2_{f1}} {N_{f1}} + \frac{SD^2_{f2}} {N_{f2}} +
 \frac{SD^2_{m1}} {N_{m1}} + \frac{SD^2_{m2}} {N_{m2}}
\right),
$$

where subscripts *f* and *m* stand for 'females' and 'males'.

Similarly, the dependent version of the sampling variance can be written as:

$$
\sigma^2_{ARRD} = \left( \frac{1} {T_2 - T_1} \right)^2 
\left( 
\frac{SD^2_{f1} + SD^2_{f2}  - 2rSD_{f1}SD_{f2}} {N_f}   +
\frac{SD^2_{m1} + SD^2_{m2}  - 2rSD_{m1}SD_{m2}} {N_m}
\right).
$$

Here we have derived sampling variances for 'temperature' *ARR* and *ARRD* for both independent and dependent cases. These formulas can also be applied to changes in other factors (e.g., salinity, pH, oxygen) by changing the temperature constant. For example, it may be possible that males and females were measured in slightly different temperatures; say females at $T_3$ and $T_4$ ($T_4 > T_3$) and male at $T_5$ and $T_6$ ($T_6 > T_5$) like below:

$$
\sigma^2_{ARRD} = \left( \frac{1} {T_4 - T_3} \right)^2 
\left( 
\frac{SD^2_{f3} + SD^2_{f4}  - 2rSD_{f3}SD_{f4}} {N_f}   +
\right) +
\left( \frac{1} {T_6 - T_5} \right)^2 
\left( 
\frac{SD^2_{m5} + SD^2_{m6}  - 2rSD_{m5}SD_{m6}} {N_m}
\right).
$$

Also, note that the difference between slopes does not need to be that of males and females. This formula can be used to compare any two treatments or biological groups.

### Controlling for unitary differences and temperature across studies

Calculating *ARR* assumes that the units for the mean difference across studies is consistent. Frequently, this will not be the case for meta-analyses that hope to synthesise a wide array of traits that vary in their units (e.g., ul/g, grams, seconds, minutes). Fortunately, *$lnRR$* and *SMD* (Table 1) already controls for unitary differences. As such, we can apply similar logic that we discuss above so that *SMD* corrects for both temperature and unitary differences. Table 1 provides the formula for *SMD*, its pooled standard deviation, and sampling variance. Just like with *ARR*, we can apply our temperature correction to the *SMD* formula as follows:

$$
SMD_{T} = \frac{\left(M_{1} - M_{2}\right)}{SD_{p}\left( T_{2} - T_{1}\right)}J
$$

Here we can see that the difference between $M_1$ and $M_2$ is standardised by the pooled SD, essentially correcting for unitary differences. Additionally, dividing SMD by the temperature difference results in further correcting the effect size by temperature. To derive the sampling variance for $SMD_T$ we can apply the same principle as we did above to derive a sampling variance as:

$$
s_{SMD_{T}}^2 = \left (\frac{N_{1} + N_{2}}{N_{1}N_{2}}  + \frac{SMD_{T}^2}{2\left( N_{1} + N_{2}\right)} \right) \left(\frac{1}{{T_{2} - T_{1}}}\right)^2
$$

### The Delta Method

Deriving sampling variances using common properties of variance has its limits. Therefore, we introduce a practical and widely applicable method to obtain approximate variance when the basic properties of variance cannot be applied. This is where the Delta method comes in. The general form of this method can be written as:

$$
\text{Var}(f(M_1)) \approx \text{Var}(M_1)* \left( 
f'(M_1)
\right)^2,
$$

where $f(M_1)$ represents the function of the random variable $M_1$, and importantly, $f'(M_1)$ is the first derivative of $f(M_1)$. Let's demonstrate this with a concrete example by re-deriving the sampling variance for $\ln{RR}$ defined as:

$$
\ln{RR} = \ln \left ( \frac{M_{1}}{M_{2}}\right ) = \ln{M_1} -\ln{M_2}.
$$

Here, $f(M_1) = \ln{M_1}$. By applying $f'(M_1) = 1/{M_1}$ (i.e. the first derivative of $lnM_1$ is $1/{M_1}$) to the delta method and using the variance's basic properties, we have:

$$
\sigma^2_{ln{RR}} \approx  \left ( \frac{SD_{1}^2}{N_{1}}\right ) * \left(\frac{1}{M_{1}} \right)^2 + \left ( \frac{SD_{2}^2}{N_{2}}\right ) * \left(\frac{1}{M_{2}} \right)^2 =   \left ( \frac{SD_{1}^2}{N_{1}M_{1}^2}\right ) + \left ( \frac{SD_{2}^2}{N_{2}M_{2}^2}\right )
$$

Note that the first term, $\frac{SD_{1}^2}{N_{1}}$ is once again the sampling variance for group 1, and we simply multiply this term by the square of the first derivative of $M_{1}$ as described by the Delta method. Readers may notice that this formula is now equivalent to the sampling variance for $ln_{RR}$ (Table 1). We can extend this sampling variance to include dependency between the groups as:

$$
\sigma^2_{ln{RR}} \approx  \left ( \frac{SD_{1}^2}{N_{1}M_{1}^2}\right ) + \left ( \frac{SD_{2}^2}{N_{2}M_{2}^2}\right ) 
- 2r\sqrt{\frac{SD_{1}^2}{N_{1}M_{1}^2}}\sqrt{\frac{SD_{2}^2}{N_{2}M_{2}^2}},
$$

where r is the correlation between $\ln{M_1}$ and $\ln{M_2}$ when organisms are measured multiple times (with r often assumed to be 0.5), while r is 0 when two sets of measurements are independent.

Finally, we note that, by using the basic properties of variance and the Delta method, one can derive sampling variance for most effect size measures. For instances, these methods allowed us to derive sampling variance for our 'new' effect sizes (e.g. $lnQ_{10}$; see the main text) and were used in @Nakagawa2015 and @Senior2020. We also refer the reader to @Nakagawa2017ab for additional details on the Delta method.

# References